Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-05 output/review_responses.tex

\textbf{Comment: This is the likely cause: a more thorough exploration
  of biotic stresses is needed : blast, nematodes, etc}\\

\textbf{Response: } As we do not have the data to explore this, this
discussion would be solely speculation. For more, see our response
above.\\

\textbf{Comment: Yes, this would be alarming....a steady decrease in
  the potential of all paddy Es, but why now after millennia of
  paddy?}\\

\textbf{Response: } As we state, more study is needed to verify and
determine the extent and cause of this (lines 209-210).\\

@@

\textbf{Comment: thats 1.25\% p.a. increase, almost noone is achieving
  that in any crop! Except when you jump up to hybrids, a one off
  jump. But your point is valid, yield progress is slowed by the
  "erosion" . I dont think the Figure 6,  however, is needed.}\\

\textbf{Response: } This is the rate of improvement which has been
claimed by the rice breeders in this system. This information has been
added to the manuscript (line 39). The figure is
helpful to distill the major mechanisms at work here to a simple
conceptual description.\\

### 2018-03-05 output/breed_paper_revised.Rnw
Final comments.

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\usepackage[shortcuts]{extdash}
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

Certainly one question in looking at this situation is, ``do new
cultivars have increased yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding cultivars with increased yield potential. However, when the
popular cultivar M-202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1}~\citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1}~\citep[14\%
MC;][]{m209}. However, in annual breeding reports and cultivar release
announcements, data based on yield performance in side-by-side cultivar
trials show that new cultivars usually out yield the older
cultivars. In other rice systems, a decrease in a cultivar's yield
performance over time has been observed~\citep{Peng1998, Peng2010,
Peng2000, Datta1995}, sometimes termed ``yield erosion''~\citep{Peng1998}. It
has been speculated that this yield decline over time is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
erosion may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in cultivars
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect the true magnitude of yield
changes over time.

@@

The CCRRF Rice Experiment Station (RES) is located near Biggs, CA
(39.4648, \==121.7342; Fig. \ref{fig:map}), and has been the central location for CCRRF's
efforts to develop improved rice cultivars adapted to CA since
1969. The California Statewide Variety Trials evaluate current and
promising cultivars at the RES and six to eight on-farm trials around
the state each year. To avoid potential complications due to site,
climate, and management differences at the farmer-managed trials
between years, we focused our investigation on the RES. The RES is
researcher-managed and plants the majority of the experimental plots
for the statewide program (approximately 3 to 4 times the number of
experimental plots compared to other on-farm locations). The climate
at the RES is Mediterranean, characterized by mild winters during
which most of the annual precipitation occurs (\Sexpr{18*24.5} to
\Sexpr{25*24.5} mm annually) and warm summers largely free of
precipitation events. Soils at the RES are classified as
Esquon-Neerdobe clays with roughly 2\% soil organic matter in the top
15cm~\citep{ssurgo}.

>>>>

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
%% \usepackage[shortcuts]{extdash}
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

Certainly one question in looking at this situation is, ``do new
cultivars have increased yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding cultivars with increased yield potential. However, when the
popular cultivar M-202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1}~\citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1}~\citep[14\%
MC;][]{m209}. However, in annual breeding reports and cultivar release
announcements, data based on yield performance in side-by-side cultivar
trials show that new cultivars usually out yield the older
cultivars, typically by 3--5\% (e.g., M-209 release description at \url{http://www.crrf.org/linked/2015annualreport.pdf}). In other rice systems, a decrease in a cultivar's yield
performance over time has been observed~\citep{Peng1998, Peng2010,
Peng2000, Datta1995}, sometimes termed ``yield erosion''~\citep{Peng1998}. It
has been speculated that this yield decline over time is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
erosion may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in cultivars
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect the true magnitude of yield
changes over time.

@@

The CCRRF Rice Experiment Station (RES) is located near Biggs, CA
(39.4648, -121.7342; Fig. \ref{fig:map}), and has been the central location for CCRRF's
efforts to develop improved rice cultivars adapted to CA since
1969. The California Statewide Variety Trials evaluate current and
promising cultivars at the RES and six to eight on-farm trials around
the state each year. To avoid potential complications due to site,
climate, and management differences at the farmer-managed trials
between years, we focused our investigation on the RES. The RES is
researcher-managed and plants the majority of the experimental plots
for the statewide program (approximately 3 to 4 times the number of
experimental plots compared to other on-farm locations). The climate
at the RES is Mediterranean, characterized by mild winters during
which most of the annual precipitation occurs (\Sexpr{18*24.5} to
\Sexpr{25*24.5} mm annually) and warm summers largely free of
precipitation events. Soils at the RES are classified as
Esquon-Neerdobe clays with roughly 2\% soil organic matter in the top
15cm~\citep{ssurgo}.
